Accelerating adaptive ultrasound algorithms using graphics processing units \vspace{0.5cm}

### 2014-03-16 ThesisJP/Background/background.tex
Now, only VR and Feild simulations sections remains. And some iterations with correcting language.

For the first adopters of general-purpose computing on GPUs (GPGPU) it soon became evident that offloading all types of computations to the GPU was not always a good idea. In most cases it is important to balance the load equally between the CPU and GPU to obtain their combined computationally power. Balancing typically means that highly parallel and computationally intense tasks should be offloaded to the GPU and serialized and memory intensive work should stay on the CPU. This paradigm of utilizing specialized cores for solving specific problems is today known as \textit{heterogeneous computing}. The specialized core can be a GPU or e.g. a field-programmable gate array (\nom{FPGA}{Field-programmable gate array}). A recent trend is that specialized accelerator cores are embedded onto the CPU. This heterogeneous processors is often referred to as an accelerated processing unit (\nom{APU}{Accelerated processing unit}). The last three generations of CPUs from Intel have all had on-chip integrated graphics, hence they can be classified as APUs. An on-chip GPU is less powerful than a high-performance discrete GPU, but valuable time is saved by not having to send data across the PCI Express bus, which connects the different parts of a computer to the motherboard. This is the reason why one of the first rules of GPGPU (with discrete GPUs) is to minimize CPU-to-GPU memory transfers.

@@

\subsection{GPUs in medical ultrasound imaging}
Ultrasound imaging systems generates an incredible amount of data. A modern system capable of 3D imaging will generate several tera bytes of data in the front end of the system during a clinical scan. The data rate from a 2D scan is "only" in order of giga bytes per second prior to IQ-demodulation. To handle these amount of data, ultrasound imaging systems have relied heavily on application-specific integrated circuits (\nom{ASIC}{Application-specific integrated circuit}) for all steps in the processing pipeline. However, as the computationally power of CPUs and GPUs have increased, ASICs, e.g. for converting data from a polar to a Cartesian grid (known as scan conversion), have gradually been converted into software \cite{Guracar2013}. One of  the last remaining ASICs in ultrasound scanners is the beamformer, whose job is to steer and focus the ultrasound beam in space. This operation involves a reduction of the data rate with a factor equal to the number of elements in the ultrasound probe. The input data rate is therefor very high. Nevertheless, real-time beamforming has recently been achieved on a GPU \cite{Song2012}, and a French company (SuperSonic Imagine) are for instance utilizing GPUs in their plane wave beamforming algorithm \cite{Tanter2014}. This is the same technique as reported implemented on a GPU by Yiu \textit{et al.} \cite{Yiu2011}. We therfore can conclude that beamforming is finally moving into software too. The largest players in the medical ultrasound business, General Electric and Philips, are also well aware of where the trend is going \cite{Thomenius2012} \cite{Metz2011}. Software beamforming opens up new possibilities for advanced processing of ultrasound data. One example is Capon beamforming which is explored in this thesis (\textbf{Paper\,I} and \textbf{II}), and which we will introduce in the next section. The increased computational power introduced by GPUs makes it also possible to include advanced 3D visualization as presented in \cite{solteszova2010multidirectional}, \textbf{Paper\,IV}, and \textbf{IX}.

\subsection{Adaptive beamforming}\label{sec:adaptbf}
Adaptive beamforming has several meanings in the literature on medical ultrasound imaging. It can for instance refer to aberration correction which is the task of applying time delays to compensate for the distortion introduced by fatty layers in the human body. However, in this thesis adaptive beamforming will refer to data-dependent apodization (or weighting) derived from a given optimization criterion. The optimization criterion which we focus on in this thesis is the minimum variance distortionless response criterion, which is also known as Capon's method \cite{Capon1969}. A beamformer which uses weights derived using Capon's method will in this thesis mainly be referred to as the \textit{Capon beamformer}. However, it is also known as the minimum variance  (\nom{MV}{Minimum variance beamformer}) beamformer or the minimum variance distortionless respons beamformer (\nom{MVDR}{Minimum variance distortionless respons beamformer (Capon)}).

@@

The Capon beamformer produces a weight vector, as introduced in (\ref{eq:das}), based on the impinging signal and a constrained optimization problem. The problem is as follow \cite{Capon1969}:
\begin{align}
&\min_{\vec{w}} E\{|z[n]|^2\} \rightarrow \min_{\vec{w}} \vec{w}^H \mat{\hat{R}} \vec{w} \label{eq:capon_optimization_criteria} \\
&\text{subjected to } \vec{w}^H\vec{a} = 1,
\end{align}
where \nom{$\mat{\hat{R}}$}{Sample covariance matrix (with regularization)} is a sample covariance matrix. Hence, the resulting weight vector minimizes the output power (the criterion) while maintaining unit gain in the steering direction $\vec{a}$ (distortionless response). The benefit of Capon beamforming is clearly seen from Fig\,\ref{fig:weights} which shows a sensitivity map (an \textit{array beampattern}) for DAS beamforming with uniform weights and the Capon beamformer. In this simulated case an interfering source is located at 20\degree{} which for the DAS beamformer with uniform weights (\ref{fig:das_weights}) will leak into the beamformer output with its power attenuated with around $13$ dB (first sidelobe). Since the Capon weights are data-dependent the method "knows" where the interfering source is located and is able to position a zero in the direction of it (Fig.\,\ref{fig:capon_weights}). Hence, the interfering source will be highly attenuated independent of its direction. Another interesting case is when the interfering source is located within the mainlobe of the array beampattern. In theory a signal will be regarded by the Capon beamformer as an interfering source as long as the signal propagation vector does not match the steering vector ($\vec{a}$) exactly. If it is not possible to move a zero to the direction of the interferer the mainlobe will get shifted and scaled if this results in minimum output power. Adaptive positioning of zeros and mainlobe shifting gives rice to the super-resolution capabilities of the Capon beamformer \cite{Synnevag2007}, which is the main reason why Capon beamforming has been highly studied in the field of medical ultrasound imaging the last decade.

To "know" the direction of interfering sources the Capon beamformer has to estimate the statistics of the impinging wave field. This is done by estimating a sample covariance matrix, which for an active broadband system can be estimated in the following way \cite{Synnevag2009}:
\begin{align}
\mat{\breve{R}}[n] = \frac{1}{N_LN_K}\sum_{n'=n-K}^{n+K} \sum_{l=0}^{N_L-1} \vec{x}_l[n']\vec{x}_l[n']^H,
\end{align}
where $K = M-L+1$\nomi{$K$}{Temporal averaging parameter (Capon)} typically is proportional to the pulse length, $N_K = 2K + 1$ \nomi{$N_K$}{Number of samples used for temporal averaging (Capon)}, $N_L = M-L+1$,\nomi{$N_L$}{Number of subarrays used for spatial averaging (Capon)}\nomi{$L$}{Subarray length (Capon)} and $\vec{x}_l$ is a $L$ long subarray $[x_l[n], \dotso x_{l+L}[n]]$. To ensure robustness against model errors and numerical stability the matrix is loaded with a diagonal factor, 
\begin{align}
\epsilon &= d*\text{trace}\{\mat{\breve{R}}\}/L\\
\mat{\hat{R}} &= \mat{\breve{R}} + \epsilon\mat{I}.
\end{align} 
Thus, a white noise component proportional to the in-coherent signal power is added to the estimate.

The solution to the constrained minimization problem in (\ref{eq:capon_optimization_criteria}) is, by the method of Lagrange multipliers \cite{VanTrees2003}, found to be
\begin{align}\label{eq:capon_weights}
\vec{w}[n] = \frac{\mat{\hat{R}}[n]^{-1}\vec{a}}{\vec{a}^H\mat{\hat{R}}[n]^{-1}\vec{a}} \in \mathbb{C}^L.
\end{align}
Since $\vec{w}$ is time (or range) dependent one matrix has to be constructed and inverted for each data vector $\vec{x}[n]$ received by the system. This is indeed computationally demanding, something which becomes clear when each step of the beamformer is closer examined. It is then found that construction of the covariance matrix has a computationally complexity of $O(L^2N_LN_K)$\nomi{O(.)}{Big O notation (algorithm complexity)}, inverting it has a computationally complexity of $O(L^3)$, and the final normalization and calculation of $\vec{w}$ is $O(L)$. For $L=M/2$, a subarray length often used in the literature, the first two steps both have a complexity of $O(L^3)$. These steps should therefore receive the same amount of attention when a fast implementation of Capon beamforming is attempted. A rough estimate of the number of FLOPS required by a Capon beamformer for real-time cardiac ultrasound imaging is then given by:
\begin{align}
\text{FLOPS}_{\text{Capon}} = L^3N_{\text{B-mode}}f_{\text{FPS}} = L^3B_wf_c.
\end{align}
Which for our example, with $L=M/2=64$, sums up to a required throughput of more than $1/2$ TFLOPS. Even though this is just 10\% of the peak throughput of modern GPUs remember that the latter where only theoretical numbers, and that the actual throughput is highly algorithm dependent. The required throughput for Capon beamforming might, for the moment, be hard to achieve in practice \cite{So2011}. As for DAS beamforming this number also scales with higher relative bandwidth, pulse frequency, and more array elements since the subarray size typically is proportional to the number of array elements. In the literature on medical ultrasound imaging we find few attempts on implementing a real-time Capon beamformer for medical ultrasound imaging. In addition to the papers in this thesis only Chen \textit{et al.} \cite{Chen2011, Chen, Chen2011a} are known to have explored GPU and FPGA implementations of the Capon beamformer for medical ultrasound imaging, however with some shortcomings  which are discussed in both \textbf{Paper I} and \textbf{II}.

To reduce the processing requirement of the Capon beamformer for medical ultrasound imaging several simplified version have been proposed \cite{Asl2012, Kim} in the literature. Among the most interesting proposals we find the low complexity data-dependent beamformer by Synnev\aa{}g \textit{et al.} \cite{Synnevag2011} and the beamspace (BS) Capon beamformer by Nilsen and Hafizovic \cite{Nilsen2009}. In medical ultrasound imaging beamspace typically refers to the polar grid that cardiac ultrasound data is located in prior to scan conversion. In combination with the Capon beamformer, beamspace refers to the wavenumber-space representation of the recorded signals, hence, the spatial Fourier transform of element data. A GPU implementation of the BS-Capon beamformer for real-time cardiac ultrasound imaging is presented in \textbf{Paper II}. The complexity of BS-Capon beamforming is dominated by the transformation of data into beamspace which has a computational complexity of $O(N_LLN_b)$, where \nom{$N_b$}{Number of selected beams in beamspace} is the number of beams selected in beamspace. The initial complexity of BS-Capon beamforming before any tricks are performed is then found to be:
\begin{align}
\text{FLOPS}_{\text{BS-Capon}} = L^2N_bB_wf_c
\end{align}
For our example BS-Capon beamforming, with $L=M/2=64$ and $N_b=3$, will require a throughput of around $30$ GFLOPS. This is a level of processing which both modern CPUs and GPUs should be able to cope with. The rationale behind using the GPU is that a highly parallel task like the Capon beamformer should be offloaded to the processor made for parallel processing. The CPU is then left to do managing and all the task which by tradition occupies it.

Capon beamforming and focused systems.

The Capon beamformer is derived for narrowband, far-field applications and has typically been applied on medical ultrasound imaging "as is" even tough medical ultrasound imaging is a broadband, near-field imaging application. Later in this section we will explain why a broadband version of the Capon beamformer will have little effect. First we will take a closer look at the details of Capon beamforming.
For the low complexity data-dependent beamformer all the shifted windows are shifted less than the mainlobe width.

Intro to adaptive beamforming. List other variants (LCA, beamspace, Eigen space etc.).

Add section about how to present data (max v.s mean etc.)

\subsection{Shift invariance}

One sentence about speckle tracking.

@@

Get section from master theses. Ray casting and opacity functions.

>>>>

For the first adopters of general-purpose computing on GPUs (GPGPU) it soon became evident that offloading all types of computations to the GPU was not always a good idea. In most cases it is important to balance the load equally between the CPU and GPU to obtain their combined computationally power. Balancing typically means that highly parallel and computationally intense tasks should be offloaded to the GPU and serialized and memory intensive work should stay on the CPU. This paradigm of utilizing specialized cores for solving specific problems is today known as \textit{heterogeneous computing}. The specialized core can be a GPU or e.g. a field-programmable gate array (\nom{FPGA}{Field-programmable gate array (processor)}). A recent trend is that specialized accelerator cores are embedded onto the CPU. This heterogeneous processors is often referred to as an accelerated processing unit (\nom{APU}{Accelerated processing unit}). The last three generations of CPUs from Intel have all had on-chip integrated graphics, hence they can be classified as APUs. An on-chip GPU is less powerful than a high-performance discrete GPU, but valuable time is saved by not having to send data across the PCI Express bus, which connects the different parts of a computer to the motherboard. This is the reason why one of the first rules of GPGPU (with discrete GPUs) is to minimize CPU-to-GPU memory transfers.

@@

\subsection{GPUs in medical ultrasound imaging}
Ultrasound imaging systems generates an incredible amount of data. A modern system capable of 3D imaging will generate several tera bytes of data in the front end of the system during a clinical scan. The data rate from a 2D scan is "only" in order of giga bytes per second prior to IQ-demodulation. To handle these amount of data, ultrasound imaging systems have relied heavily on application-specific integrated circuits (\nom{ASIC}{Application-specific integrated circuit}) for all steps in the processing pipeline. However, as the computationally power of CPUs and GPUs have increased, ASICs, e.g. for converting data from a polar to a Cartesian grid (known as scan conversion), have gradually been converted into software \cite{Guracar2013}. One of  the last remaining ASICs in ultrasound scanners is the beamformer, whose job is to steer and focus the ultrasound beam in space. This operation involves a reduction of the data rate with a factor equal to the number of elements in the ultrasound probe. The input data rate is therefor very high. Nevertheless, real-time beamforming has recently been achieved on a GPU \cite{Song2012}, and a French company (SuperSonic Imagine) are for instance utilizing GPUs in their plane wave beamforming algorithm \cite{Tanter2014}. This is the same technique as reported implemented on a GPU by Yiu \textit{et al.} \cite{Yiu2011}. We therfore can conclude that beamforming is finally moving into software too. The largest players in the medical ultrasound business, General Electric and Philips, are also well aware of where the trend is going \cite{Thomenius2012} \cite{Metz2011}. Software beamforming opens up new possibilities for advanced processing of ultrasound data. One example is Capon beamforming which is explored in this thesis (\textbf{Paper\,I} and \textbf{II}), and which is introduce in the next section. The increased computational power introduced by GPUs makes it also possible to include more advanced 3D visualization as presented in \cite{solteszova2010multidirectional}, \textbf{Paper\,IV}, and \textbf{IX}.

\subsection{Adaptive beamforming}\label{sec:adaptbf}
Adaptive beamforming has several meanings in the literature on medical ultrasound imaging. It can for instance refer to aberration correction which is the task of applying time delays to compensate for the distortion introduced by fatty layers in the human body. However, in this thesis adaptive beamforming will refer to data-dependent apodization (or weighting) derived from a given optimization criterion. The optimization criterion which we focus on in this thesis is the minimum variance distortionless response criterion, which is also known as Capon's method \cite{Capon1969}. A beamformer which uses weights derived using Capon's method will in this thesis mainly be referred to as the \textit{Capon beamformer} \cite{Vignon2008}. However, it is also known as the minimum variance  (\nom{MV}{Minimum variance beamformer (Synonym for Capon)}) beamformer \cite{Synnevag2007}, the minimum variance distortionless respons beamformer (\nom{MVDR}{Minimum variance distortionless respons beamformer (Synonym for Capon)}) (\textbf{Paper I}), or the constrained adaptive beamformer \cite{Mann2002}.

@@

The Capon beamformer produces a weight vector, as introduced in (\ref{eq:das}), based on the impinging signal and a constrained optimization problem. The problem is as follow \cite{Capon1969}:
\begin{align}
&\min_{\vec{w}} E\{|z[n]|^2\} \rightarrow \min_{\vec{w}} \vec{w}^H \mat{\hat{R}} \vec{w} \label{eq:capon_optimization_criteria} \\
&\text{subjected to } \vec{w}^H\vec{a} = 1,
\end{align}
where \nom{$\mat{\hat{R}}$}{Sample covariance matrix (with regularization)} is a sample covariance matrix, and \nom{$(.)^H$}{Complex conjugate transpose} is the complex conjugate transpose. Hence, the resulting weight vector minimizes the output power (the criterion) while maintaining unit gain in the steering direction $\vec{a}$ (distortionless response). The benefit of Capon beamforming is clearly seen from Fig\,\ref{fig:weights} which shows a sensitivity map (an \textit{array beampattern}) for DAS beamforming with uniform weights and the Capon beamformer. In this simulated case an interfering source is located at 20\degree{} which for the DAS beamformer with uniform weights (\ref{fig:das_weights}) will leak into the beamformer output with its power attenuated with around $13$ dB (first sidelobe). Since the Capon weights are data-dependent the method "knows" where the interfering source is located and is able to position a zero in the direction of it (Fig.\,\ref{fig:capon_weights}). Hence, the interfering source will be highly attenuated independent of its direction. Another interesting case is when the interfering source is located within the mainlobe of the array beampattern. In theory a signal will be regarded by the Capon beamformer as an interfering source as long as the signal propagation vector does not match the steering vector ($\vec{a}$) exactly. If it is not possible to position a zero in the direction of the interferer the mainlobe will get shifted and scaled if this is whats needed to achieve minimum output power. Adaptive positioning of zeros and mainlobe shifting gives rice to the super-resolution capabilities of the Capon beamformer \cite{Synnevag2007}, and the positive implications of this effect \cite{Synnevag2009} is the main reason why this beamformer has been highly studied in the field of medical ultrasound imaging the last decade.

To "know" the direction of interfering sources the Capon beamformer has to estimate the statistics of the impinging wave field. This is done by estimating a sample covariance matrix, which for an active broadband system can be estimated in the following way \cite{Synnevag2009}:
\begin{align}
\mat{\breve{R}}[n] = \frac{1}{N_LN_K}\sum_{n'=n-K}^{n+K} \sum_{l=0}^{N_L-1} \vec{x}_l[n']\vec{x}_l[n']^H,
\end{align}
where\nomi{$K$}{Temporal averaging parameter (Capon)} $N_K = 2K + 1$ \nomi{$N_K$}{Number of samples used for temporal averaging (Capon)} is the number of samples used for temporal smoothing,\nomi{$L$}{Subarray length (Capon)} $\vec{x}_l$ is a $L$ long subarray $[x_l[n], \dotso, x_{l+L}[n]]$, and $N_L = M-L+1$\nomi{$N_L$}{Number of subarrays used for spatial averaging (Capon)} is the number of subarrays used for spatial averaging.  Temporal averaging is used to obtain DAS-like speckle statistics in combination with large subarrays \cite{Synnevag2007a}, and subarray averaging is applied to reduce signal cancellation caused by coherent signals \cite{Reddy1987}. The number of temporal samples is usually selected relative to the pulse length. To ensure robustness against model errors and numerical stability the matrix is further loaded with a diagonal factor, 
\begin{align}
\epsilon &= d*\text{trace}\{\mat{\breve{R}}\}/L\\
\mat{\hat{R}} &= \mat{\breve{R}} + \epsilon\mat{I}.
\end{align} 
Hence, a white noise component proportional to the in-coherent signal power is added to the estimate \cite{Featherstone1997b}.

The solution to the constrained minimization problem in (\ref{eq:capon_optimization_criteria}) is, by the method of Lagrange multipliers \cite{VanTrees2003}, found to be
\begin{align}\label{eq:capon_weights}
\vec{w}[n] = \frac{\mat{\hat{R}}[n]^{-1}\vec{a}}{\vec{a}^H\mat{\hat{R}}[n]^{-1}\vec{a}} \in \mathbb{C}^L.
\end{align}
Since $\vec{w}$ is time (or range) dependent one matrix has to be constructed and inverted for each data vector $\vec{x}[n]$ received by the system. This is indeed computationally demanding, something which becomes clear when each step of the beamformer is examined closer. It is then found that constructing the covariance matrix has a computationally complexity of $O(L^2N_LN_K)$\nomi{O(.)}{Big O notation (algorithm complexity)}, inverting it has a computationally complexity of $O(L^3)$, and the final normalization and calculation of $\vec{w}$ is $O(L)$. For $L=M/2$, a subarray size often used in the literature, the first two steps both have a complexity of $O(L^3)$. These steps should therefore receive the same amount of attention when a real-time implementation of Capon beamforming is attempted.

A rough estimate of the number of FLOPS required by a Capon beamformer for real-time cardiac ultrasound imaging is then given by:
\begin{align}
\text{FLOPS}_{\text{Capon}} = L^3B_wf_c.
\end{align}
Which for our example, with $L=M/2=64$, sums up to a required throughput of more than $1/2$ TFLOPS. Even though this is just 10\% of the peak throughput of modern GPUs remember that the latter are only theoretical numbers, and that the actual throughput is highly algorithm dependent. The required throughput for Capon beamforming might therefore still be hard to achieve in practice \cite[\textbf{Paper II}]{So2011}. As for DAS beamforming this number also scales with higher relative bandwidth, pulse frequency, and more array elements since the subarray size typically is proportional to the number of array elements.

In the literature on medical ultrasound imaging we find few attempts on implementing a real-time Capon beamformer for medical ultrasound imaging. In addition to the papers in this thesis only Chen \textit{et al.} \cite{Chen2011, Chen, Chen2011a} are known to have explored GPU and FPGA implementations of the Capon beamformer for medical ultrasound imaging. The differences between our work and Chen \textit{et al.} is discussed in \textbf{Paper I}, \textbf{II}, and \textbf{VI}.

To reduce the processing requirement of the Capon beamformer for medical ultrasound imaging several simplified version have been proposed in the literature \cite{Asl2012, Kim}. Among the most interesting proposals we find the low complexity adaptive (LCA) beamformer by Synnev\aa{}g \textit{et al.} \cite{Synnevag2011} and the beamspace (BS) Capon beamformer by Nilsen and Hafizovic \cite{Nilsen2009}. In medical ultrasound imaging beamspace typically refers to the polar grid that cardiac ultrasound data is located in prior to scan conversion. In combination with the Capon beamformer, beamspace refers to the wavenumber-space representation of the recorded signals, hence, the spatial Fourier transform of the element data: 
\begin{align}
&\mat{B}[x_l[n],\,\dotso,\,x_{l+L}[n]]^T\\
&[\mat{B}]_{p,q} = \frac{1}{\sqrt{L}}e^{-j 2 \pi p q / L}
\end{align}
The idea behind BS-Capon beamforming is that spatial beams with little information can be removed from $\mat{\hat{R}}$ by removing rows in $B$. The complexity of BS-Capon beamforming is dominated by the transformation of data into beamspace which has a computational complexity of $O(N_LLN_b)$, where \nom{$N_b$}{Number of selected beams in beamspace} is the number of selected beams.

The number of FLOPS required for BS-Capon beamforming of real-time cardiac ultrasound is then found to be:
\begin{align}
\text{FLOPS}_{\text{BS-Capon}} = L^2N_bB_wf_c
\end{align}
For our example, BS-Capon beamforming with $L=M/2=64$ and $N_b=3$, will require a throughput of around $30$ GFLOPS. This is a level of processing which both modern CPUs and GPUs should be able to handle. A GPU implementation of the BS-Capon beamformer for real-time cardiac ultrasound imaging is presented in \textbf{Paper II}. The rationale behind using the GPU is that a highly parallel task like the Capon and BS-Capon beamformer should be offloaded to the processor made for parallel processing (heterogeneous computing). The CPU is then left to do managing and all the task which by tradition already occupies it. It should also be possible to achieve higher processing throughput on a GPU than on a CPU for this kind of problem.

Both the LCA and BS-Capon beamformer is based upon a detailed investigation of how Capon beamfoming behaves when applied to medical ultrasound images. For the LCA beamformer, the windows selected by the Capon beamformer where investigated and a special set of predefined windows where defined. If one look closer at the windows selected for the LCA beamformer we see that half the set performs micro-steering of the mainlobe and the other half provides different mainlobe widths. The results from applying the LCA beamformer shows that the main benefit of applying Capon beamforming to medical ultrasound imaging is micro-steering of the mainlobe, and not attenuation of far off-axis backscatter.

The Capon beamformer is derived for narrowband applications and has traditionally been applied on medical ultrasound imaging "as is". Medical ultrasound imaging is, however, a broadband imaging modality and one could think that a broadband Capon beamformer therefore where needed \cite{Holfort2009}. The investigation performed in \cite{Nilsen2009} shows that with focused beams, energy will be centered on-axis and as little as three near-axis beams is therefore adequate in order to achieve similar performance for BS-Capon and Capon beamforming. The two-way beamprofile will be wider with un-focused (plane wave) transmit beams \cite{Holfort2009, Austeng2011}, but focusing on receive will still make the main amount of energy arrive on axis. As it is shown in \textbf{Paper III} it is possible to micro-steer with phase delays instead of time delays without introducing large errors as long as the phase shift is smaller than the pulse length. Since Capon beamforming mainly performs main-lobe shifts smaller than the beam width, broadband Capon beamforming will only provide a slight improvement over narrowband Capon beamforming.

%A final important point when working with Capon beamforming is how we present our results with respect to normalization. Since the Capon beamforming highly attenuate signals within the mainlobe  we can get a false impression of improvements in contrast if we normalize  \cite[Fig.\,6]{Synnevag2007}
%\todo{Need figures here...}

%Examples of other adaptive apodization methods found in the literature on medical ultrasound imaging includes the APES beamformer, the MUSIC method, and the eigenvalue method. \todo{Need references}.

%\subsection{Shift invariance}

%One sentence about speckle tracking.

@@

Get section from master theses. Ray casting and opacity functions. (Fig.\,\ref{fig:vr})
